---
title: "Prediction Competion 1"
date: "15/01/2024"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include=FALSE}
library(ggplot2)
```
Anonymized name: Sukuna

```{r, include=FALSE}
education <- c(11.28278, 13.68312, 14.40669, 15.96969, 14.86386, 15.33683, 12.69982, 11.447, 11.87571, 11.87264, 15.14703, 11.12808, 15.11679, 12.33917, 16.24666, 16.42583, 15.69144, 12.7121, 14.64092, 10.02925)
income_est <- c(142500, 165000, 180000, 200001, 190000, 195000, 161000, 145000, 148000, 147500, 191000, 140000, 190500, 152000, 210000, 215000, 197500, 162000, 182000, 130000)

data = data.frame(education, income_est)

write.table(data, "/Users/andrew/Downloads/UW courses/ECON 626/Prediction Competition 1/pc1.csv", sep=",",row.names=FALSE, col.names=FALSE)
```
## Question 2a

```{r}
model = lm(income_est ~ education)
model_summary = summary(model)
b0 = model_summary$coefficients[1]
b1 = model_summary$coefficients[2]
print(model_summary)
```


```{r, echo=FALSE, fig.height=3, fig.width=6}
ggplot(data, aes(x=education, y=income_est)) +
  labs(x = "Years of Education", y = "Estimated Income") +
  labs(title = "Estimated incomes based on years of education") +
  theme(plot.title = element_text(hjust = 0.5)) +
  geom_point() +
  geom_abline(intercept = b0, slope = b1, color = "red", linetype = "dashed")
```

```{r}
y1 = round(b0 + b1*(12))
y2 = round(b0 + b1*(13))

income_diff = round(b1 * (13 - 12))
```

\$`r sprintf("%.0f", income_diff)`.

In this economy, I estimate Jimmys income with 12 years of education to be \$`r sprintf("%.0f", y1)` and his income with 13 years of education to be \$`r sprintf("%.0f", y2)`. The difference in expected lifetime earnings for 13 years of education compared to 12 years is \$`r sprintf("%.0f", income_diff)`. I arrived at this answer using my linear regression model.

## Question 2b

```{r}
current_avg_income = mean(income_est)
future_avg_income = b0 + b1*(17)

avg_income_diff = future_avg_income - current_avg_income
```

\$`r sprintf("%.0f",avg_income_diff)`

In this economy I estimate the average lifetime earnings increases by \$`r sprintf("%.0f",avg_income_diff)` if the government assigned everyone 17 years of education. I arrived at this answer through computing the average of my predictions incomes, computing my linear models average at 17 years, then taking the difference of the two averages.

## Question 3

a) The measured college outcomes are first-year cumulative GPA, percentage of grades that are A or A- and an indicator for academic struggle defined as having any grade of C+ or lower in the first year.

b) The variables that are the most important predictors of college outcomes a student’s standardized test score. The least important predictor of college outcomes are a student’s high school GPA.


c) In my opinion, the student’s standardized test score is the most important variable. I believe this because all students are meant to take the same standardized exam and as stated in the paper, a persons socioeconomic background has little effect on the SAT and how well the individual does in their first year of college.

## Question 4

I asked chat gpt:

can you turn these two columns into two variables in R script [11.28278 & 1425000 \\
 13.68312 & 165000\\
 14.40669 & 180000\\
 15.96969 & 200000\\
 14.86386 & 190000\\
 15.33683 & 195000\\
 12.69982 & 161000\\
 11.447 & 145000\\
 11.87571 & 148000\\
 11.87264 & 147500\\
 15.14703 & 191000\\
 11.12808 & 140000\\
 15.11679 & 190500\\
 12.33917 & 152000\\
 16.24666 & 210000\\
 16.42583 & 215000\\
 15.69144 & 197500\\
 12.7121 & 162000\\
 14.64092 & 182000\\
 10.02925 & 130000\\]

Chatgpt replied:

Certainly! In R, you can create two variables from the given data using the c() function. Here's an example of how you can do it:


data <- c(11.28278, 13.68312, 14.40669, 15.96969, 14.86386, 15.33683, 12.69982, 11.447, 11.87571, 11.87264, 15.14703, 11.12808, 15.11679, 12.33917, 16.24666, 16.42583, 15.69144, 12.7121, 14.64092, 10.02925)
labels <- c(1425000, 165000, 180000, 200000, 190000, 195000, 161000, 145000, 148000, 147500, 191000, 140000, 190500, 152000, 210000, 215000, 197500, 162000, 182000, 130000)


variable1 <- data
variable2 <- labels


print(variable1)
print(variable2)

